Hydras and Cie \\ Variations on Kirby \& Paris' hydra battles and ordinal numbers in Coq

### 2021-09-20 doc/JFLA2022/Ex42c.tex
small corrections

\begin{alectryon}
  % Generator: Alectryon
  \sep
  \begin{sentence}
    \begin{input}
      ~~\PY{n+nb}{assert}~\PY{o}{(}\PY{n}{HAP}\PY{o}{:}~\PY{n}{In}~\PY{n}{AP}~\PY{o}{(}\PY{n}{phi0}~\PY{l+m+mi}{2}\PY{o}{))}~\PY{n+nb+bp}{by}~\PY{n+nb}{apply}~\PY{n}{AP\PYZus{}phi0}\PY{o}{.}\nl
    \end{input}
  \end{sentence}
  \sep
  \begin{sentence}
    \begin{input}
      ~~\PY{n+nb}{elim}~~\PY{n}{HAP}\PY{o}{;}~\PY{n+nb}{intros}~~\PY{n}{\PYZus{}}~\PY{n}{H0}\PY{o}{;}~\PY{n+nb}{apply}~\PY{n}{H0}\PY{o}{;}~\PY{n+nb}{clear}~\PY{n}{H0}\PY{o}{.}
    \end{input}
    \sep
    \begin{output}
      \begin{goals}
        \begin{goal}
          \begin{hyps}
            \hyp{HAP}{\PY{n}{In}~\PY{n}{AP}~\PY{o}{(}\PY{n}{phi0}~\PY{o}{(}\PY{n}{F}~\PY{l+m+mi}{2}\PY{o}{))}}
          \end{hyps}
          \sep
          \infrule{}
          \sep
          \begin{conclusion}
            \PY{n+nb+bp}{omega}~\PY{o}{+}~\PY{n}{F}~\PY{l+m+mi}{42}~\PY{o}{\PYZlt{}}~\PY{n}{phi0}~\PY{o}{(}\PY{n}{F}~\PY{l+m+mi}{2}\PY{o}{)}
          \end{conclusion}
        \end{goal}
      \end{goals}
    \end{output}
  \end{sentence}
\end{alectryon}

>>>>

\begin{alectryon}
  % Generator: Alectryon
  \sep
  \begin{sentence}
    \begin{input}
      ~~\PY{n+nb}{assert}~\PY{o}{(}\PY{n}{Hlt}\PY{o}{:}~\PY{n+nb+bp}{omega}~\PY{o}{\PYZlt{}}~\PY{n}{phi0}~\PY{o}{(}\PY{n}{F}~\PY{l+m+mi}{2}\PY{o}{))}~\PY{n+nb+bp}{by}~\nl
      ~~~~~~\PY{o}{(}~\PY{n+nb}{rewrite}~\PY{n}{omega\PYZus{}eqn}\PY{o}{;}~\PY{n+nb}{apply}~\PY{n}{phi0\PYZus{}mono}\PY{o}{,}~\PY{n}{finite\PYZus{}mono}\PY{o}{;}\nl
      ~~~~~~~~\PY{n+nb}{auto~with}~\PY{n}{arith}\PY{o}{).}
    \end{input}
    \sep
    \begin{output}
      \begin{goals}
        \begin{goal}
          \begin{hyps}
            \hyp{HAP}{\PY{n}{In}~\PY{n}{AP}~\PY{o}{(}\PY{n}{phi0}~\PY{o}{(}\PY{n}{F}~\PY{l+m+mi}{2}\PY{o}{))}}
            \sep
            \hyp{Hlt}{\PY{n+nb+bp}{omega}~\PY{o}{\PYZlt{}}~\PY{n}{phi0}~\PY{o}{(}\PY{n}{F}~\PY{l+m+mi}{2}\PY{o}{)}}
          \end{hyps}
          \sep
          \infrule{}
          \sep
          \begin{conclusion}
            \PY{n+nb+bp}{omega}~\PY{o}{+}~\PY{n}{F}~\PY{l+m+mi}{42}~\PY{o}{\PYZlt{}}~\PY{n}{phi0}~\PY{o}{(}\PY{n}{F}~\PY{l+m+mi}{2}\PY{o}{)}
          \end{conclusion}
        \end{goal}
      \end{goals}
    \end{output}
  \end{sentence}
\end{alectryon}

### 2021-09-20 doc/JFLA2022/paper.tex
change the Makefile
small corrections

%%% \snippets{A,B,C,…} inputs a series of snippets as one block (with \itemsep
%%% between them).  A, B, C should be paths to files in movies/snippets/.
\usepackage{etoolbox}
\makeatletter
\newcommand{\inputsnippets}[1]
  {{\setlength{\itemsep}{1pt}\setlength{\parsep}{0pt}% Adjust spacing
    \alectryon@copymacros\begin{io}
      \forcsvlist{\item\@inputsnippet}{#1}
    \end{io}}}
\let\input@old\input % Save definition of \input
\newcommand{\@inputsnippet}[1]
  {{\renewenvironment{alectryon}{}{}% Skip \begin{alectryon} included in snippet
    \input@old{../movies/snippets/#1}}}
\makeatother

@@

\begin{abstract}
  \textbf{To do!} We present the current state of a \textit{Coq} library on ordinal numbers.

\end{abstract}

@@

%------------------------------------------------------------------------------
\section{Introduction}
\label{sect:introduction}
Hydra games (also known as \emph{Hydra battles}) appear in an article published in 1982 by two mathematicians, 
Laurie Kirby and Jeff Paris: \emph{Accessible Independence Results for Peano Arithmetic}~\cite{KP82}.
This article describes a game between two players: Hercules and a hydra.
A short description  can be found in Andrej Bauer's blogpost~\cite{bauer2008} as well as a simulator~\cite{BauerHydra}.
In a few words:
\begin{itemize}
\item A hydra is a finite tree, traditionnaly presented with the root at the bottom, the leaves of which are called \emph{heads}
  (Fig~\vref{fig:iota-example}).
\item At every round, Hercules chops off one head of the hydra. If the head is at a distance greater or equal than 2 from the root,
  then some sub-tree $h$ of the hydra is copied a certain amount $i$ of times. The number $i$ of copies and the copied sub-hydra $h$ may depend of the considered variant of the game
  or the time elapsed since the beginning of the fight.
\end{itemize}
   \begin{figure}[htb]
\centering
\begin{tikzpicture}[very thick, scale=0.3]
\node (foot) at (10,0) {$\bullet$};
\node (N1) at (2,2) {$\bullet$};
\node (N2) at (10,2) {$\bullet$};
\node (N22) at (7,2) {$\bullet$};
\node (N3) at (14,2) {$\bullet$};
\node (N4) at (18,2) {$\Smiley[2][green]$};
\node (N5) at (0,4) {$\bullet$};
\node (N6) at (2,5) {$\Smiley[2][green]$};
\node (N7) at (4,6) {$\Smiley[2][green]$};
\node (N88) at (7,4) {$\bullet$};
\node (N8) at (10,4) {$\bullet$};
\node (N9) at (14,6) {$\Smiley[2][green]$};
\node (N10) at (0,8) {$\Smiley[2][green]$};
\node (N11) at (10,7) {$\Smiley[2][green]$};
\node (N111) at (7,7) {$\Smiley[2][green]$};
\draw (foot) to [bend left=10] (N1);
\draw (foot) -- (N2);
\draw (foot) -- (N22);
\draw (foot) -- (N3);
\draw (foot) -- (N4);
\draw (N1) to  (N5);
\draw (N1) to   [bend left=10] (N6);
\draw (N1) to   [bend right=20] (N7);
\draw (N2) to  (N8);
\draw (N22) to  (N88);
\draw (N8) to  (N11);
\draw (N88) to  (N111);
\draw (N3) to  (N9);
\draw (N5) to  (N10);
\end{tikzpicture}
\caption{The hydra associated with $\omega^{\omega+2}+\omega^\omega \times 2 + \omega + 1$ \label{fig:iota-example}}

@@

\begin{figure}[h]
  \centering
  \begin{itemize}
\item If $\alpha=0$, then $H'_\alpha (i)= i$ for any natural number $i$.
\item If $\alpha=\beta+1$, then 
$H'_\alpha(i)=H'_\beta(i+1)$ for any $i \in \mathbb{N}$.
\item If $\alpha$ is a limit ordinal, then 
$H'_\alpha(i) = H'_{(\canonseq{\alpha}{i+1})}(k)$ for any $i\in \mathbb{N}$.
\end{itemize}
\vspace{4pt}
\inputsnippets{Hprime/HprimeDef}
\caption{The rapidly growing hierarchy $H'$ of arithmetic functions}
\label{fig:Hprime}
\end{figure}

@@

  \begin{figure}[h]
    \centering
    \fbox{
      \begin{minipage}[h]{1.0\linewidth}
        \inputsnippets{E0/Ex42}
      \end{minipage}}
    \caption{A simple proof by computation}
    \label{fig:Ex42E0}
  \end{figure}

@@

\inputsnippets{Schutte/Ex42a}

@@

\inputsnippets{Schutte/Ex42b}

@@

\inputsnippets{Schutte/Ex42d, Schutte/Ex42c, Schutte/Ex42e}

@@

>>>>

%%% \snippets{A,B,C,…} inputs a series of snippets as one block (with \itemsep
%%% between them).  A, B, C should be paths to files in movies/snippets/.
\usepackage{etoolbox}
\makeatletter
\newcommand{\inputsnippets}[1]
  {{\setlength{\itemsep}{1pt}\setlength{\parsep}{0pt}% Adjust spacing
    \alectryon@copymacros\begin{io}
      \forcsvlist{\item\@inputsnippet}{#1}
    \end{io}}}
\let\input@old\input % Save definition of \input
\newcommand{\@inputsnippet}[1]
  {{\renewenvironment{alectryon}{}{}% Skip \begin{alectryon} included in snippet
    \input@old{#1}}}
\makeatother

@@

\begin{abstract}
  \textbf{To do!} 
\end{abstract}

@@

%------------------------------------------------------------------------------
\section{Introduction}
\label{sect:introduction}
The \emph{Hydra-battles} project, part of \emph{Coq-Community}~\cite{CoqCommunity}  aims to be an experimental platform for the collaborative development of commented libraries of formal proofs.

Hydra games (also known as \emph{Hydra battles}) appear in an article published in 1982 by two mathematicians, 
Laurie Kirby and Jeff Paris: \emph{Accessible Independence Results for Peano Arithmetic}~\cite{KP82}.
This article describes a game between two players: Hercules and a hydra.
A short description  can be found in Andrej Bauer's blogpost~\cite{bauer2008} as well as a simulator~\cite{BauerHydra}.
In a few words:
\begin{itemize}
\item A hydra is a finite tree, traditionnaly presented with the root at the bottom, the leaves of which are called \emph{heads}
  (Fig~\vref{fig:iota-example}).
\item At every round, Hercules chops off one head of the hydra. If the head is at a distance greater or equal than 2 from the root,
  then some sub-tree $h$ of the hydra is copied a certain amount $i$ of times. The number $i$ of copies and the copied sub-hydra $h$ may depend of the considered variant of the game
  or the time elapsed since the beginning of the fight.
\end{itemize}
   \begin{figure}[htb]
\centering
\begin{tikzpicture}[very thick, scale=0.3]
\node (foot) at (10,0) {$\bullet$};
\node (N1) at (2,2) {$\bullet$};
\node (N2) at (10,2) {$\bullet$};
\node (N22) at (7,2) {$\bullet$};
\node (N3) at (14,2) {$\bullet$};
\node (N4) at (18,2) {$\Smiley[2][green]$};
\node (N5) at (0,4) {$\bullet$};
\node (N6) at (2,5) {$\Smiley[2][green]$};
\node (N7) at (4,6) {$\Smiley[2][green]$};
\node (N88) at (7,4) {$\bullet$};
\node (N8) at (10,4) {$\bullet$};
\node (N9) at (14,6) {$\Smiley[2][green]$};
\node (N10) at (0,8) {$\Smiley[2][green]$};
\node (N11) at (10,7) {$\Smiley[2][green]$};
\node (N111) at (7,7) {$\Smiley[2][green]$};
\draw (foot) to [bend left=10] (N1);
\draw (foot) -- (N2);
\draw (foot) -- (N22);
\draw (foot) -- (N3);
\draw (foot) -- (N4);
\draw (N1) to  (N5);
\draw (N1) to   [bend left=10] (N6);
\draw (N1) to   [bend right=20] (N7);
\draw (N2) to  (N8);
\draw (N22) to  (N88);
\draw (N8) to  (N11);
\draw (N88) to  (N111);
\draw (N3) to  (N9);
\draw (N5) to  (N10);
\end{tikzpicture}
\caption{The hydra associated with $\omega^{\omega+2}+\omega^\omega \times 2 + \omega + 1$ \label{fig:iota-example}}

@@

\begin{figure}[h]
  \centering
  \begin{itemize}
\item If $\alpha=0$, then $H'_\alpha (i)= i$ for any natural number $i$.
\item If $\alpha=\beta+1$, then 
$H'_\alpha(i)=H'_\beta(i+1)$ for any $i \in \mathbb{N}$.
\item If $\alpha$ is a limit ordinal, then 
$H'_\alpha(i) = H'_{(\canonseq{\alpha}{i+1})}(k)$ for any $i\in \mathbb{N}$.
\end{itemize}
\vspace{4pt}
\inputsnippets{HprimeDef}
\caption{The rapidly growing hierarchy $H'$ of arithmetic functions}
\label{fig:Hprime}
\end{figure}

@@

  \begin{figure}[h]
    \centering
    \fbox{
      \begin{minipage}[h]{1.0\linewidth}
        \inputsnippets{Ex42}
      \end{minipage}}
    \caption{A simple proof by computation}
    \label{fig:Ex42E0}
  \end{figure}

@@

\inputsnippets{Ex42a}

@@

\inputsnippets{Ex42b}

@@

\inputsnippets{Ex42d, Ex42c, Ex42e}

@@

 \emph{Hydra-battles} is not limited to the study of ordinal numbers and applications.
 For instance, we are also developing 
a module about efficient exponentiation algorithms, and hope
to extend our project to new topics.
